### 2015-06-08 LOI/inc2.tex
still need to finesse this better
Refocused technical approach to a storage focus.

The I/O and storage system prototype will enable
highly flexible data access modes that are common in scientific applications
and associated analysis.
%
%\hasan{Can we actually take some lower accuracy representation of data (low
%  granularity MR for example, and compute more stuff to provide a higher
%  accuracy representation? I am thinking of the work with Varis Carey with
%  storing boundary state instead of full volume and then using those
%  boundaries to fill in blanks over temporally changing data}
%
\textbf{First}, data re-computation is allowed to re-generate data with
customizable accuracy, if the associated computing and I/O cost is lower
than retrieving the target data from a {\em slow} tier.
%
\textbf{Second}, exploratory analysis operations frequently entail an
overall data set view followed by targeted data exploration on extracted
features.  In particular, the {\em overview} access mode gives a quick but
approximate data view that can guide feature selection offering rapid
coarse-grained data exploration without requiring loading the high-accuracy
data from storage. Based on the granularity requested, the accuracy and size
of the data returned can be adjusted. In the most extreme setting, the
original data can be retrieved at the time cost of moving the potentially
huge data quantity.
%
\textbf{Third}, to ensure available storage for subsequent operations, we
will offer automatic data migration, based on user annotations for required data and using monitoring and learning techniques.  Unlike existing approaches, this will be
driven both by the user annotations and through learned access patterns in
an autonomic and crosslayer manner.  While past access patterns may not
indicate future access because the simulation run purpose may have changed,
we are focused on scalability where subsequent runs are larger as the
simulation prepares for a capability run.  By learning from the output and
access patterns during this run sequence, we can accurately decide how to
place and organize data for the critical capability runs.
%
\textbf{Fourth}, we anticipate storing multiple data copies, each compressed
in different ways according to the underlying media, some of these copies
will disappear based on storage pressures, but data persistence will be
maintained according to user specifications. Assuming a relatively low
latency cache layer before a tape system, we can offer exploratory data
access reserving pulling data from tape to just the data required. 
 Additionally, we will incorporate data life time, at sub-object
granularity, as a user described metric to allow the storage system to
intelligently make decisions on purging and evicting data.

>>>>

The I/O and storage system prototype will enable
highly flexible data access modes that are common in scientific applications
and associated analysis.
%
%\hasan{Can we actually take some lower accuracy representation of data (low
%  granularity MR for example, and compute more stuff to provide a higher
%  accuracy representation? I am thinking of the work with Varis Carey with
%  storing boundary state instead of full volume and then using those
%  boundaries to fill in blanks over temporally changing data}
%
\textbf{First}, the storage system will offer the ability to directly store
different portions of a single dataset--even a single variable at differnt
levels of the storage hierarchy and apply different data transformation
operations to each portion. For example, full data can be queued to tape while
a highly compressed version intended for high-level analytical views can be
stored in NVM or even RAM within the storage hierarchy. The storage system will
support the metadata and plug-ins reuired to support this sort of data storage
approach.
%data re-computation is allowed to re-generate data with
%customizable accuracy, if the associated computing and I/O cost is lower
%than retrieving the target data from a {\em slow} tier.
%
\textbf{Second}, new storage access APIs will be developed that understand the
notion of a time and data quality for requested data. The storage system,
through the knowledge of both the data quality stored in differeht storage
hierarchy tiers and the approximate time to retrieve data from the different
tiers, the user can manage the tradeoff between data precision and time to
retrieve. This storage system support will be managed through middleware
insulating the user as much as possible from the new APIs. The storage system
must also support plug-ins to potentially decompress or expand data to the
original size. Both lossless and lossy with error bounds style storage is
assumed. The time factor must take into account how long this operation takes
for servicing the user request.
%exploratory analysis operations frequently entail an
%overall data set view followed by targeted data exploration on extracted
%features.  In particular, the {\em overview} access mode gives a quick but
%approximate data view that can guide feature selection offering rapid
%coarse-grained data exploration without requiring loading the high-accuracy
%data from storage. Based on the granularity requested, the accuracy and size
%of the data returned can be adjusted. In the most extreme setting, the
%original data can be retrieved at the time cost of moving the potentially
%huge data quantity.
%
\textbf{Third}, the storage system will offer annotations within the metadata
and support for both predictive and reactive data placement and migration.
While past access patterns may not indicate future access because the
simulation run purpose may have changed, we are focused on scalability where
subsequent runs are larger as the simulation prepares for a capability run.  By
learning from the output and access patterns during this run sequence, we can
accurately decide how to place and organize data for the critical capability
runs. The reactive mechanisms will consider the space and performance
characteristics, as well as the requested data fidelity, to determine where and
how to pull and store data sets. In some cases, a full fidelity will be pulled
from tape to disk and no further to preserve 100\% data fidelity. In others, a
lossy-compressed within error bounds version may be generated and stored in
local node RAM for very fast access with an acceptable error.
%to ensure available storage for subsequent operations, we
%will offer automatic data migration, based on user annotations for required
%data and using monitoring and learning techniques.  Unlike existing
%approaches, this will be driven both by the user annotations and through
%learned access patterns in
%an autonomic and crosslayer manner.
%
\textbf{Fourth}, to effectively manage data storage capacities, the storage
system's capability for storing multiple, potentially different data quality
versions of the same data will be leveraged to offer eviction and migration. The main challenge of this approach is to keep in mind the intentional placement
decisions made to optimize future data access while ensuring proper system
operation by not exhausting any tier inappropriately. If a single application
has been allocated the entire machine for a capability run, completely
consuming resources on the machine is expected. For smaller runs, these
allocations must be balanced to support the diverse workload.
%we anticipate storing multiple data copies, each compressed
%in different ways according to the underlying media, some of these copies
%will disappear based on storage pressures, but data persistence will be
%maintained according to user specifications. Assuming a relatively low
%latency cache layer before a tape system, we can offer exploratory data
%access reserving pulling data from tape to just the data required. 
% Additionally, we will incorporate data life time, at sub-object
%granularity, as a user described metric to allow the storage system to
%intelligently make decisions on purging and evicting data.
